A modern, event-based architecture for distributed evolutionary algorithms

### 2018-04-03 paper.tex
Down to 13

  There is usually no direct translation of most algorithms to this
  architecture. This is specially true in evolutionary algorithms, who
  rely on state, in the shape of a population, to perform their search
  in the space of solutions.
  In this paper we want to introduce KafkEO, a serverless framework for
  evolutionary algorithms and other population-based systems. The main
  design objective is to leverage the scaling capabilities of a
  serverless framework, as well as create a system that can be deployed
  on different platforms by using free software. Our intention has
  also been to create an algorithm that is functionally equivalent to an asynchronous
  parallel, island-based, EA, which can use parallelism and at the same
  time reproduce mechanisms that are akin to migration. The island-based 
  paradigm is relatively standard in distributed EA applications, but in our case, 
  we have been using it since it allows for better parallelism
  and thus performance, at the same time it makes keeping diversity
  easier while needing fewer parameters to tune.

@@

  In general, scientific computing has followed the trends of the
  computing industry, with new implementations published almost as soon
  as new technologies became commercially available, or even
  before. There were very early implementations of evolutionary
  algorithms on transputers \cite{voigt1990modelling}, the world wide
  % transputers, had to google that
  % you're too young, man - JJ
  web \cite{chong:1999:jDGPi} or the first generation of cloud
  services
  \cite{DBLP:journals/corr/abs-1105-6205,de2017parallel,salza2017ccube}. Evolutionary
  algorithms are population-based algorithms, so it is relatively easy
  to create a parallel implementation by just splitting populations, as
  long as there is a way for those populations to interchange some
  individuals and leverage the use of many different machines. In
  general, besides, one of the most time-consuming steps in an
  evolutionary algorithm is evaluation, so as long as you can compute
  fitness in parallel, you can obtain a certain amount of speed up.

  However, in general every new computing platform has its own view of
  how computing should be made, and in many cases that has made
  evolutionary algorithms move in new directions in order to make them
  work better in that platform while keeping the spirit of
  bioinspiration. For instance, most evolutionary algorithms work in a
  synchronous way; although there were very early efforts to create
  asynchronous EAs \cite{coleman89}, in general
  generations proceed one after the other and migration happening in
  all islands at the same time. However, this mode of working does not
  fit well with architectures that are heterogeneous and dynamic, which
  is why there have been many efforts from early
  on to adapt EAs to this kind of substrate
  \cite{Jini:FEA2000,zorman2002creation,baugh2003asynchronous}. This
  last one creates an asynchronous, parallel system that is {\em functionally
    equivalent} to a sequential one by {\em unrolling} operations so
  that they happen in parallel as long as there is no dependence between
  them; multiple generations occur at the same time with every operation
  using data that is available at every particular
  moment. \cite{Jini:FEA2000} and \cite{zorman2002creation} explicitly
  follow two completely different, and opposed, approaches to the
  adaptation of an EA to a heterogeneous substrate. While the former
  tends towards finer grain by distributing every particular operation
  to different Jini services, the latter uses a pool-based approach to
  create island-like structures, creating again a system that, even
  leveraging the capabilities of the architecture in a much better way,
  does not advance the  structure of an EA in any new way.

@@

  In general, frameworks based in SoA tried to achieve functional equivalence with parallel
  or sequential versions of EAs. There is the same tension between
  functional equivalence and new
  design in new, cloud based approaches to evolutionary
  algorithms. Salza and collaborators \cite{salza2017ccube,de2017parallel} explicitly
  and looking to optimize interoperability claim that there is very
  little need to change ``traditional'' code to port it to the cloud,
  implicitly claiming this {\em functional equivalence} with the
  sequential evolutionary algorithms. Other contemporary systems such as
  \cite{10.1007/978-3-319-32149-3_46} use Spark to evaluate, in
  parallel, fitness for different machine learning models;
  Vianna de Oliveira and Coelho \cite{de2015scalable} use Hadoop for evolutionary k-means clustering,
  obtaining good speedups; in this last case, it uses a problem-specific
  way of distributing tasks, without resourcing to a change of the
  evolutionary algorithm itself.

  Besides these implementations using well known cloud services, there are new computation models for evolutionary algorithms
  that are not functionally equivalent to a canonical EA, but have
  proved to work well in these new environments. Pool based EAs,
  \cite{bollini1999distributed}, with a persistent population that can
  be tapped to retrieve single individuals or pools of them and return
  evaluated or evolved sub-populations, have been used for new
  frameworks such as EvoSpace \cite{García-Valdez2015}, and proved to be
  able to accommodate all kinds of ephemeral and heterogeneous
  resources.

  In the serverless, event based types of architectures we are going to
  be targeting in this paper, there has been so far no work that we know
  of. Similar setups including microservices have been employed by Salza et
  al. \cite{salza2017ccube}; however, the serverless system adds a layer
  of abstraction to event-based queuing systems such as the one employed
  by Salza by reducing it to functions, messages and rules or
  triggers. We will explain in detail these architectures in the next
  section.

>>>>

  %There is usually no direct translation of most algorithms to this
  %architecture. This is specially true in evolutionary algorithms, who
  %rely on state, in the shape of a population, to perform their search
  %in the space of solutions.
  In this paper we want to introduce KafkEO, a serverless framework for
  evolutionary algorithms and other population-based systems. The main
  design objective is to leverage the scaling capabilities of a
  serverless framework, as well as create a system that can be deployed
  on different platforms by using free software. Our intention has
  also been to create an algorithm that is functionally equivalent to an
  %% Pool Based ??
   asynchronous, parallel, island-based, EA, which can use parallelism and at the same
  time reproduce mechanisms that are akin to migration. The island-based 
  paradigm is relatively standard in distributed EA applications, but in our case, 
  we have been using it since it allows for better parallelism
  and thus performance, at the same time it makes keeping diversity
  easier while needing fewer parameters to tune.

@@

  In general, scientific computing has followed the trends of the
  computing industry, with new implementations published almost as soon
  as new technologies became commercially available, or even
  before. There were very early implementations of evolutionary
  algorithms on transputers \cite{voigt1990modelling}, the world wide
  % transputers, had to google that
  % you're too young, man - JJ
  web \cite{chong:1999:jDGPi} and the first generation of cloud
  services
  \cite{DBLP:journals/corr/abs-1105-6205,de2017parallel,salza2017ccube}. %Evolutionary
  %algorithms are population-based algorithms, so it is relatively easy
  %to create a parallel implementation by just splitting populations, as
  %long as there is a way for those populations to interchange some
  %individuals and leverage the use of many different machines. In
  %general, besides, one of the most time-consuming steps in an
  %evolutionary algorithm is evaluation, so as long as you can compute
  %fitness in parallel, you can obtain a certain amount of speed up.
  However, every new computing platform has its own view of computing, and in many cases that has made evolutionary algorithms move in new directions in order to make them work better in that platform while keeping the spirit of
  bio-inspiration. For instance, most evolutionary algorithms work in a
  synchronous way; although there were very early efforts to create
  asynchronous EAs \cite{coleman89}, in general
  generations proceed one after the other and migration happens in
  all islands at the same time. However, this mode of working does not
  fit well with architectures that are heterogeneous and dynamic, which
  is why there have been many efforts from early
  on to adapt EAs to this kind of substrate
  \cite{Jini:FEA2000,zorman2002creation,baugh2003asynchronous}.
  %This last one creates an asynchronous, parallel system that is {\em functionally equivalent} to a sequential one by {\em unrolling} operations so
  %that they happen in parallel as long as there is no dependence between
  %them; multiple generations occur at the same time with every operation
  %using data that is available at every particular moment.
  %\cite{Jini:FEA2000} and \cite{zorman2002creation} explicitly
  %follow two completely different, and opposed, approaches to the
  %adaptation of an EA to a heterogeneous substrate. While the former
  %tends towards finer grain by distributing every particular operation
  %to different Jini services, the latter uses a pool-based approach to
  %create island-like structures, creating again a system that, even
  %leveraging the capabilities of the architecture in a much better way,
  %does not advance the  structure of an EA in any new way.

@@

  In general, frameworks based in SoA also tried to achieve functional equivalence with parallel
  or sequential versions of EAs. There is the same tension between
  functional equivalence and new
  design in new, cloud based approaches to evolutionary
  algorithms. 
  %TODO: The previous should be improved, is very important. 
  Salza and collaborators \cite{salza2017ccube,de2017parallel} explicitly
  and looking to optimize interoperability claim that there is very
  little need to change ``traditional'' code to port it to the cloud,
  implicitly claiming this {\em functional equivalence} with
  sequential evolutionary algorithms.
  % Not that important or related? - Mario
  %Other contemporary systems such as
  %\cite{10.1007/978-3-319-32149-3_46} use Spark to evaluate, in
  %parallel, fitness for different machine learning models;
  %Vianna de Oliveira and Coelho \cite{de2015scalable} use Hadoop for evolutionary k-means clustering,
  %obtaining good speedups; in this last case, it uses a problem-specific
  %way of distributing tasks, without resourcing to a change of the
  %evolutionary algorithm itself.

  % Yes! we got to the chase 
  Besides these implementations using well known cloud services, there are new computation models for evolutionary algorithms
  that are not functionally equivalent to a canonical EA, but have
  proved to work well in these new environments. Pool based EAs,
  \cite{bollini1999distributed}, with a persistent population that can
  be tapped to retrieve single individuals or pools of them and return
  evaluated or evolved sub-populations, have been used for new
  frameworks such as EvoSpace \cite{García-Valdez2015}, and proved to be
  able to accommodate all kinds of ephemeral and heterogeneous
  resources.

  In the serverless, event-based architectures we are going to
  be targeting in this paper, there has been so far no work that we know
  of. Similar setups including microservices have been employed by Salza et
  al. \cite{salza2017ccube}; however, the proposed serverless system adds a layer of abstraction to event-based queuing systems such as the one employed
  by Salza by reducing it to functions, messages and rules or
  triggers. We will explain in detail these architectures in the next
  section.
